---
author: "Fernando Mateos-González"
date: Last update "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    code_download: true
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
  md_document:
    variant: markdown_github
  pdf_document:
    toc: yes
title: "Capture-Recapture Analysis for Miroslav Susliks 2020"
editor_options: 
  markdown: 
    wrap: 72
---

# Introduction

This report analyses data from 2 capture-recapture sessions in Miroslav
during the 2020 spring in order to calculate the density of the
population.

Session 1 run from 23,27 of April to May 1st (6 occasions).

Session 2 run from 18 - 22 of May (5 occasions).

(Note: No problem with different n of occasions according to Murray
<http://www.phidot.org/forum/viewtopic.php?f=36&t=3260>)

```{r global-options, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE, results="hide"}

knitr::opts_chunk$set(fig.width=6, fig.asp = 0.618, fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE)

options(allow_html_in_all_outputs=TRUE)
```

# Libraries

```{r results="hide"}
library(secr)
library(tidyverse)
library(here)

```

# Tidying data

Saved and edited two txt files from the original data directly in excel
(sacrilege!): captures and traps.

-   Eliminated several columns that we won't use in this analysis

-   Selected only captures in the traps used for CR.

-   Deleted one capture of a female without collar, a single capture in
    between sessions, and two captures without collar code.

-   Recoded the sessions, which were coded as every single day, into
    sessions 1 and 2. (within each session we can consider the
    population as closed, but deaths and births can happen between
    sessions).

-   Renamed the variable "Unique ID" as "occasion" and recoded it - now,
    correctly - as each day as an occasion.

-   Renamed the Collar ID variable as captureid. We'll use the collar
    code this time, but we need an individual unique code for each
    animal across seasons and hopefully across years if the collars stay
    on the susliks after hibernation.

-   Used the same trap layout from last year and added character 0 to
    single digit traps ids on the capture file to match the trap file

# Analysis

We combine the information obtained from our captures and the trap
locations to create the capthist file. The analysis was performed using
the package 'secr' version 4.3.3 and R `r getRversion()`.

```{r capthist}

multi <- read.capthist(here("data", "captures.txt"), here("data", "traps.txt"), detector = "proximity") #The “proximity” detector type allows, at
# most, one detection of each individual at a particular detector on any occasion (i.e.,
# it is equivalent to what we call the Bernoulli or binomial encounter process model,
# or model SCR0).

summary(multi)


```

*n number of distinct individuals detected on each occasion t\
*u number of individuals detected for the first time on each occasion t\
*f number of individuals detected on exactly t occasions\
*M(t+1) cumulative number of detected individuals on each occasion t

## Trap array and consecutive captures of each individual

```{r}
par(mfrow = c(1,1), mar = c(1,1,1,1)) # reduce margins
plot (multi, tracks = TRUE, gridsp = 5,border = 10)
      

```

The most important insight from these figures is that individuals tend
to be recaptured near their site of first capture. This is expected when
the individuals of a species occupy home ranges. In SECR models the
tendency for detections to be localised is reflected in the spatial
scale parameter σ.

## Successive trap-revealed movements and distances 

```{r}
m <- unlist(moves(multi))
par(mar = c(3.2,4,1,1), mgp = c(2.1,0.6,0)) # reduce margins
hist(m, breaks = seq(0/5, 50,5), xlab = "Movement m", main = "")
```

We will employ the estimate of the spatial scale σ from 2019
(6.21) to fit the simplest possible SECR model with function secr.fit.

```{r}
fit <- secr.fit (multi, buffer = 4 * 6.21, trace = FALSE,biasLimit = NA, verify = FALSE)

detector(traps(multi)) <- "proximity"

fit
```

The report comprises:\
• function call and time stamp\
• summary of the data\
• description of the model, including the maximized log likelihood,
Akaike's Information Criterion AIC\
• estimates of model coefficients (beta parameters)\
• estimates of variance-covariance matrix of the coefficients\
• estimates of the 'real' parameters

. The estimated density is 158.58 susliks per hectare, 95% confidence
interval 102-246 susliks per hectare\
. The other two real parameters jointly determine the detection
function, plotted below with 95% confidence limits

```{r}
par(mar = c(4,4,1,1)) # reduce margins
plot(fit, limits = TRUE)


```

The theory of SECR tells us that buffer width is not critical as long as
it is wide enough that animals at the edge have effectively zero chance
of appearing in our sample.We check that for the present model with the
function esa.plot.

The estimated density has easily reached a plateau at the chosen buffer
width (dashed red line):

```{r}
esa.plot(fit)
abline(v = 4 * 6.21, lty = 2, col = 'red')

```

## Homeranges

Mean distance in meters, between consecutive capture locations, pooled
over individuals, per capture session (e.g. Efford 2004).

```{r }
dbar(multi, userdist = NULL, mask = NULL)

```

MMDM (for 'Mean Maximum Distance Moved') is the average maximum distance
(in m) between detections of each individual i.e. the observed range
length averaged over individuals (Otis et al. 1978).

```{r}

MMDM(multi, min.recapt = 1, full = FALSE, userdist = NULL, mask = NULL)

```

RPSV (for 'Root Pooled Spatial Variance') is a measure of the 2-D
dispersion of the locations at which individual animals are detected,
pooled over individuals (cf Calhoun and Casby 1958, Slade and Swihart
1983).

```{r}
RPSV(multi, CC = FALSE)
```
